\documentclass[12pt,letterpaper,noanswers]{exam}
%\usepackage{color}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage[margin=0.9in]{geometry}
\renewcommand{\familydefault}{\sfdefault}
\usepackage{multicol}
\pagestyle{head}
\definecolor{c03}{HTML}{FFDDDD}
\header{AM 108 Final Project}{}{Due Friday \\ Dec 18th}
\runningheadrule
\headrule
\usepackage{diagbox}
\usepackage{graphicx} % more modern
%\usepackage{subfigure} 
\usepackage{amsmath} 
\usepackage{amssymb} 
%\usepackage{gensymb} 
%\usepackage{natbib}
\usepackage{hyperref}
%\usepackage{enumitem}
%\setlength{\parindent}{0pt}
%\usepackage{setspace}
%\pagestyle{empty}  
%\newcommand{\Sc}[0]{
%{\color{BlueViolet}\S}
%}
\usepackage{tcolorbox}
\usepackage[framed,numbered,autolinebreaks,useliterate]{mcode}

\begin{document}
 \pdfpageheight 11in 
  \pdfpagewidth 8.5in
  
\textbf{Project Assessment Information}


\begin{enumerate}
    \item 
Aims of the project:
\begin{itemize}
    \item What is the math being used to do (can I tell)?
    \item If the mathematical effort were successful what would be useful about having done that math (what questions could you answer with this math)?
    \item Does it make sense to pursue this project in this course?  (How good a fit is the topic of the project with the content of this course?)
\end{itemize}

\item
Project content:
\begin{itemize}
    \item Are all of the elements of the project clearly connected together?
    \item Does the scope of the mathematical effort surpass the mathematical effort involved in a problem set?
    \item Is the mathematics clearly linked to the content of the course?
    \item Is the mathematics that is presented correct?
    \item  Do the mathematical arguments presented contain and/or abstract away enough details that they could be followed by classmates?
    \item How original is the mathematical effort?
    \item How much of an extension on the course material (in terms of numerical tools, math methods, scope, or depth), if at all, is the mathematical effort?
\end{itemize}


\item
Presentations:
\begin{itemize}
    \item Is there a clear story that connects the entire presentation?
    \item  Is the mathematics discussed in the presentation presented in a way that it is understandable for students in the class?
    \item Is the mathematical effort discussed in the presentation connected back to the aims of the project?
    \item Are assumptions and limitations of the work clearly spelled out (and are the key ones captured)?
    \item Is time used well in the presentation (with adequate time dedicated to the mathematical results)?
    \item Are graphics used to illustrates content and support audience understanding (and appropriately formatted)?
    \item Do presenters avoid reading from slides? (no content on the slides is serving as presenter notes)
\end{itemize}
% \item
% Report:
% \begin{itemize}
% \item In development.
% \end{itemize}
\end{enumerate}

\textbf{Components of the final presentation}
\begin{itemize}
\item question/goal
\item background your classmates will need to understand your work
\item approach/model
\item data sources (if relevant)
\item results/conclusions
\item recommendations for next steps (what you would do if you had more time)
\item references
\end{itemize}

\section{The individual report}
\begin{enumerate}
    \item Write the aim / goal of your group's project work in your own words (this should be written completely independently of other members of your team).
    \item Submit your work log for Dec 6th - 18th.
    \item Briefly summarize your contributions to your group's project work.
    \item If there are aspects of your work that were not visible in the presentation and were not included in your progress reports, describe those here.  
    % \begin{itemize}
    % \item Describe how your work on the project was reflected in the presentation.  Document your contributions.
    %     \item Describe any reading, modeling, or other exploration that you pursued and that is not reflected in the two presentations.
        
    %     \emph{This question is about your work on the project, not the overall work of your team}.
    %     \item Submit any code you have contributed to, summaries or notes you may have written, or other work products you created as part of your project work that you would like to share with me.  For joint work that you submit, make sure to inform me about how you contributed.
    %     \item If you have continued to keep a log of your working time since the project update as part of problem set 10, submit that log as part of this report.
    % \end{itemize}
\end{enumerate}
\end{document}
